Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2021-02-06 paper/paper.tex
paper: Update Abstract
Update Abstract with changes suggested by @cguz

Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
\author{CD. Álvaro, CAG. Álvarez, J. Álvaro}
\date{24th of January, 2021}

@@

\resumen{
The characterization and understanding of \emph{Open Clusters}
allow us to better understand properties and mechanisms about the Universe
such as stellar formation and the regions where these events occur.
They also provide information about stellar processes and the evolution of the galactic disk.

In this paper, we develop a method to characterize open clusters by using
\emph{machine learning} tools, such as clustering by \emph{K-Means}
and clustering based on \emph{Artificial Neural Networks} by adapting
the \emph{Deep Embedded Clustering} model.
We are using \emph{Gaia DR2 database} as data source.

The developed method aims to improve classic techniques of the state of the art.
We achieve improvements not only in terms of \emph{computational efficiency},
with lower computational requirements, but in \emph{usability}, reducing the number
of hyperparameters to get a good characterization of the analyzed clusters.

Our method achieves good results,
becoming even better in some cases when the results are compared with current techniques.
}

>>>>

%---------------------------
%título del trabajo y autor
%---------------------------
\title{Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data}
\author{CD. Álvaro, C. Guzman, J. Álvaro}
\date{24th of January, 2021}

@@

\resumen{
The characterization and understanding of \emph{Open Clusters} (OCs)
allow us to understand better properties and mechanisms about the Universe
such as stellar formation and the regions where these events occur.
They also provide information about stellar processes and the evolution
of the galactic disk.

In this paper, we present a novel method to characterize OCs.
Our method employs a model built on \emph{Artificial Neural Networks} (ANNs).
More specifically, we adapted a state of the art model, the
\emph{Deep Embedded Clustering} (DEC) model for our purpose.

The developed method aims to improve classical state of the arts techniques.
We improved not only in terms of computational efficiency, (with lower
computational requirements), but in usability (reducing the number
of hyperparameters to get a good characterization of the analyzed clusters).
For our experiments, we used the \emph{Gaia DR2 database} as the data source,
and compared our model with the clustering technique \emph{K-Means}.

Our method achieves good results,
becoming even better (in some of the cases) than current techniques.
}
